### 2022-09-09 project-timeline.rmd
detailed project timeline with dates for deadlines

# Project timeline

## Week 1 (29/9 - 2/9)
- Setup Github ([done](https://github.com/victoria-l-w/trauma-severity-ofi))
- Draft study progress report
- Reading

## Week 2 (5/9 - 9/9) 
- Continue drafting study progress report
- Detailed project  timeline

## Week 3 (12/9 - 16/9)
- Complete draft of study progress report so that there's a week+ for feedback
- Start writing code that imports and begins to clean data
- Start drafting a more publication-friendly version

## Week 4 (19/9 - 23/9)
- Revise study progress report based on feedback, submit
- Revise code based on feedback.
- Add code that creates a first table of sample characteristics.
- Continue improving publication-friendly version

## Week 5-10 (26/9 - 4/11)
- **Deadline**: Submit study progress report by 12:00 Monday 26/9 
- **Presentations**: Present study progress report 4-5/10 
- Write code that generate a first set of preliminary results and
  revise weekly based on feedback.

## Week 11-12 (7/11 - 18/11)
- Draft the Results section and upload.
- Revise code based on feedback.

## Week 13-14 (21/11 - 2/12)
- Draft Discussion section and upload.
- Revise Results based on feedback.

## Week 15 (5/12 - 9/12)
- Draft and upload final project paper.

## Week 16 (12/12 - 16/12) 
- Revise final project paper and upload a second version.

## Week 17 (19/12 - 23/12)
- Revise second version of final project paper and submit.

## Week 18-20 (26/12 - 13/1)
- **Deadline**: Submit seminar version by 12:00 Tuesday 27/12
- **Presentations**: Seminar presentations either 5/1 or 9/1
- **Deadline**: Submit final version by 12:00 Friday 13/1
- Revise project paper based on comments and feedback from peers,
  course leaders, examiners, and supervisors