### 2022-09-12 meetingnotes.rmd
meeting notes 12/9

## 12/9
**Attending**: Me, George, Jonatan

**Discussed**:

* Using RISCII is probably not possible because the database lacks necessary parameters. Just use TRISS and NORMIT, or find a new third (or more?) score to use? Something for me to look into and think about. 
* Meeting KI requirements and deadlines more important than aims to publish, KI is prio 1.

**Plan**

* Access database
* Review what's available in the database + look into scoring systems, decide if there are any suitable scoring systems that would be interesting to add -> this would mean I'd have to calculate any extra scores myself b/c they aren't ready-made in the database
* Complete draft of projektredogörelse by end of week

**Next meeting**: Tuesday 20/12 over lunch, either Zoom or irl.