### 2022-09-12 studyprogressreport.Rmd
lots of text about m&m; other small fixes

---
title: "Association between trauma severity models and opportunities for improvement: A retrospective cohort study"
author: "Victoria Witthoff"
output:
  pdf_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Abbreviations
=============
* ASA-PS - American Society of Anesthesiology Physical Status
* KUH - Karolinska University Hospital
* KUH TCQD - Karolinska University Hospital Trauma Care Quality Database
* KUH TR - Karolinska University Hospital Trauma Registry
* NORMIT - Norwegian Survival Prediction Model in Trauma
* OFI - Opportunities for improvement
* Ps - Probability of survival
* TRISS - Trauma and Injury Severity Score

Introduction
============
<!-- Why trauma is a big deal -->

Traumatic injury is one of the leading causes of death worldwide, constituting nearly 8% of global deaths annually [@whoepi]. The global health burden of physical injury has led to trauma being described as a worldwide pandemic. One of the characteristics of this particular pandemic is its demographic spread; trauma is a leading cause of death for people aged 10-49 years [@globalburden] and 90% of trauma-related deaths occur in low- to middle-income countries [@whoepi]. Nonfatal outcomes also comprise a significant portion of the global health burden of disability [@globalburden], which can drastically reduce quality of life and health status in survivors of severe trauma [@globalburden; @longtermtrauma].

Management of severe trauma is highly time-sensitive and is also dependent on the actions of a multidisciplinary chain of healthcare providers. Errors in management are a common cause of preventable deaths [@prevdeathsghorbani; @prevdeathsmontmany; @prevdeathsteixeira]. Outcomes in severe trauma are dependent on the quality of care received, and it has been shown in the US that trauma quality improvement (QI) programs improve outcomes [@ustqi]. However, it has been shown that the implementation of QI programs differs widely between clinics [@tqipanglo] and there is still plenty of progress to be made in improving these programs.

## Morbidity and mortality review

The gold standard of trauma quality improvement programs worldwide are multidisciplinary morbidity and mortality (M&M) conferences.  M&M conferences review selected patient cases and attempt to establish whether there were opportunities for improvement (OFI) in each specific case, and, where necessary, implement corrective actions. The incidence of OFI function in this manner as a proxy for incidence of suboptimal care. The WHO guidelines for trauma quality improvement programmes recommend regularly scheduled multidisciplinary conferences and give specific recommendations as to their structure and format [@whoguidelines]. Despite these  recommendations, there is a still a great degree of variation amongst clinics in how M&M conferences are carried out, with a significant number of clinics still using an informal and unstructured format [@tqpianglo]. The WHO note themselves in their guidelines that M&M conferences are not used to their fullest potential globally [@whoguidelines].

The value of M&M conferences in long-term quality improvement is dependent on selection of appropriate cases for review. The process by which patient cases are selected for M&M review differs between clinics and case selection is often non-standardised [@mmreview; @tqipanglo]. This weakness of M&M review has been subject to considerable efforts to develop strategies to standardise case selection. The benefit of standardising case selection is twofold; it could increase identification of the cases that are most likely to have been mismanaged, and it could also save resources in the form of work hours. The latter benefit is not insignificant considering the large caseload found in most urban trauma centres. Some clinics have already made efforts to standardise case election; for example by using quality indicators, or audit filters, to flag cases for inclusion in M&M conferences [@tqipanglo]. Examples of quality indicators include time to first medical contact, adverse event rate, and mortality rate [@developqi]. The WHO recommends that all "deaths, complications, adverse events and errors" should be reviewed [@whoguidelines].

<!-- Explain the what and why of TSM -->

## Trauma severity scores
Trauma severity scores (TSS) are used in trauma quality improvement programs and research for their usefulness in benchmarking and in comparing trauma severity between patients [@whoguidelines]. The existence of these models is necessitated by the heterogeneous nature of trauma, due to the variety in its underlying causes, which affect both the type and severity of the injury. This heterogeneity poses challenges in comparing and evaluating trauma outcomes and has spurred the development of multiple TSS. TSS were initially intended for use in prehospital triage, and therefore were simplistic by design for ease of use [@traumareview], but over the years newer and more advanced models have been developed for quality assurance and research purposes. Today there are hundreds of TSS models. These models differ in their methodology but all have the aim of translating injury severity into quantitative values.

One of the limitations of any TSS model is that their estimates may be unreliable when applied to a population different from the one the model was developed from.

The Trauma and Injury Severity Score (TRISS), first developed in 1987 [@triss] and most recently updated in 2010 <!-- Cite 2009 revision ref -->, is one of the most commonly used models worldwide.TRISS is a combination index that incorporates patient age, injury severity by anatomical site, and physiological parameters [@triss]. The resulting value is a percentage that estimates the probability of survival (Ps).

<!-- Triss Limitations -->

<!-- NORMIT -->

<!-- Explain why TSM ⇒ M&M -->

 <!--Kolla exv Strömmers och Ghorbanis pek,https://pubmed-ncbi-nlm-nih-gov.sll.idm.oclc.org/29797712/ .
De försöker använda bl.aa normit och triss för att se korrelans till OFI och preventable deaths. Skillnaden mot det du gör är att de bara tittar på patienter som dör och inte hela cohorten. Du tittar också på ett betydligt större dataunderlag. Så ditt bygger vidare på deras arbete-->

 However, the extent to which TSS are associated with or can predict OFI as found in M&M conferences is poorly understood. If it is found that an existing TSS model is associated with OFI, it could allow for another avenue of automatically identifying potentially mismanaged cases in a standardised fashion. This study aims to evaluate how trauma severity models are associated with and to what extent they can predict opportunities for improvement.
<!--Introduktionen är mycket bra! Egentligen bara fortsätta som du gör, läs på och lägg till meningar/källor när du hittar info. För att nå KI:s sidkrav kan du skriva en hel del om de olika Scorsen -->

Methods 
=======

Study design
------------

<!-- Present key elements of study design -->
This is a single-centre retrospective cohort study, using data from two registries at Karolinska University Hospital (KUH): the trauma registry and the trauma care quality database (TCQD). The TCQD is a subset of the trauma registry and contains data on trauma patients who were selected for M&M reviews. These two databases will be linked and then analysis will be performed to estimate the association between the common TSS models TRISS and NORMIT <!-- And something else? --> and OFI using logistic regression. The predictive performance of these models will be evaluated using measures of discrimination and calibration.

Setting
-------

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->

KUH is located in Stockholm, Sweden. KUH's trauma centre is equivalent to a level 1 trauma centre according to American College of Surgeons standards [@nksstats]. KUH is the primary trauma centre for the region of Stockholm. In addition, KUH has agreements with several other regions to provide trauma care for patients with severe trauma. KUH is thereby the major trauma centre for a population of almost 3 million people [@nksstats].

The trauma registry at KUH is also submitted to SweTrau, Sweden's national trauma registry database, and therefore meets SweTrau guidelines. SweTrau follows the Utstein Trauma Template for reporting of data following major trauma. The registered data includes prehospital vital signs and management, vital signs and management on arrival at hospital, information about the type of injury, and the primary method of treatment. Follow-up variables include, but are not limited to, survival at 30 days, days in ventilator, GCS at discharge, highest level of care, and transfers to other units.

The trauma care quality register is an internal hospital register and registers the outcome of M&M conferences. Each case reviewed is registered in the trauma care quality register with data about the outcome of the review.

Participants
------------

Participants include all patients registered in both the trauma registry and the TCQD at KUH. <!--När du kommit igång med analysen så ska faktiskt nummer in här (yes ska göra det) -->

### Selection for the trauma registry

The trauma registry includes all patients for whom the trauma team was activated after receiving a potentially traumatic injury, regardless of NISS score, and all patients with a NISS score >15, regardless of whether the trauma team was activated. Also included are patients who were transferred to the hospital within 7 days of injury and have a NISS score >15. Patients where the only traumatic injury is a chronic subdural hematoma and patients for whom the trauma team was activated without an underlying traumatic injury are excluded from the register.

### Selection for the TCQD

The TCQD contains patients from the trauma registry that were selected for M&M review. At KUH, all fatalities are automatically selected for M&M review. In addition to this, the electronic medical records of all trauma patients are screened by research nursing staff to identify cases that potentially received sub-optimal care. These nurses also use audit filters that automatically highlight cases with abnormal parameters. Cases that two research nurses have evaluated and found to have potential for sub-optimal care are also selected for M&M review.

Variables
---------

<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
### Outcome
The studied outcome is the binary variable "opportunities for improvement" (OFI), as identified by the multidisciplinary M&M conference at KUH. OFI is coded as either "Yes - at least one opportunity for improvement" or "No - no opportunities for improvement".

### Exposures
The primary exposures of interest are trauma severity scores according to the TRISS and NORMIT models. Both of these models estimate the Ps based on expected mortality within 30 days. This is expressed a percentage and is a continuous variable. The variables integrated into these models are taken from arrival at hospital.

#### TRISS

The TRISS model incorporates patient age, Injury Severity Score (ISS), Revised Trauma Score (RTS), and whether the trauma was blunt or penetrating.

The ISS is given by estimating the severity of physical injuries by anatomical region. Severity is assessed on a seven-point scale from "No injury" to "Unsurvivable injury". The value given is points on a scale from 0 to 75. The RTS is based on physiological parameters GCS, systolic blood pressure, and respiratory rate. Values are taken from arrival at hospital. These parameters are weighted and expressed in a score from 0 to 8, where a lower score indicates more severe injury.

#### NORMIT

The NORMIT model incorporates the patient's age, New Injury Severity Score (NISS), comorbities prior to traumatic injury based on the ASA physical status classification system (ASA-PS), and their RTS. The NISS differs from the ISS in that the NISS is the sum of the squares from only the three most injured anatomical regions.

Bias 
----

<!-- Describe any efforts to address potential sources of bias -->

The method and data analysis model will be written using simulated data to reduce the risk of research bias. Real-world data will only be used once the data analysis model is completed and found to work correctly on simulated data.

Study size 
----------

<!-- Explain how the study size was arrived at -->

Since all eligible participants must exist in both the trauma and the trauma care quality register, and the trauma care quality registry is a subset of the trauma registry, the number of participants is limited to the number of cases registered in the trauma care quality database. The trauma care quality database includes around 2200 patients who were selected for review between 2014 and 2021.

<!-- Update these numbers later when I have real numbers -->

Quantitative variables
----------------------
<!--Om du behöver "skapa" ny data genom att manipulera den insamlade (tvek i ditt fall) så ska du beskriva det här. Annars kan detta stycke slås ihop med variables ovan. -->
<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->

### Exposures

<!--Om du inte tydligt definerat modellernas variabler behöver du göra det här, alltså för exv triss som är GCS, RR, BT + ISS osv osv-->
<!-- Really not sure exactly what to put here
-->

Statistical methods
-------------------

R, a programming language for statistical computing, will be used for compilation and analysis of data. Conversion and handling of variables will be carried out according to the SweTrau manual. Results will be presented with a confidence interval of 95%. A p-value of <0.05 will be considered significant. 
<!-- Jo det kommer nog bli ccs, du kommer exkludera alla patienter som delsa saknar outcome, OFI alternativt där det saknas nog data för att bygga olika scores. -->
<!-- Complete case study? -->
<!--

Should answer these questions, where applicable:

* All statistical methods, incl. those used to control for confounding
* Any methods to examine subgroups and interactions
* How was missing data addressed?
* Cohort study: If applicable, explain how loss to follow-up was addressed 
* Any sensitivity analyses? -->

Ethical considerations
----------------------

All patient information used in this study is anonymised and stored in a secure database. Patient medical records will not be accessed. Patients were not asked to give consent for inclusion in the KUH trauma registry. The ability and ethical approval to collect patient data for this registry without pre-emptively asking for patient consent is well-established, as collecting this data is deemed to be in the public interest according to article 6.1 e of GDPR regulations. Collecting data is in the public interest due to the fact that this patient register is used to improve management of trauma, which is a leading cause of death. Patients have the right to be removed from the register if they wish, and also have the right to a copy of data that is stored about them on request.

This project is a retrospective cohort study and the project has thereby had no effect on patient treatment. There is no risk that patients are harmed by inclusion in the study. This study has the potential to improve patient care by analysing a potential factor that could predict suboptimal trauma care. It is true that the patient population for this study includes vulnerable populations, such as the deceased and the severely injured, but this is justified by the fact that this population stands to gain the most from improved trauma care.

**Stockholm Research Ethics Review Board approval number 2021-02541 and 2021-03531.**

<!--

source("main.R")  "imports" the main script file of your project and run any code in it -->

Development and time plan
==========

<!--
B) En skriftlig redogörelse på ca en A4-sida av projektets utveckling, plus tidsplan för
resterande del av kursen (använd tidsplanmallen i Canvas). Ange så exakt som möjligt:
• Vilken typ av data, kategorier, variabler etc du avser att samla in
• Hur data planeras presenteras (t ex tabeller och hur de kan struktureras, diagramtyper,
vävnadssnitt, intervjucitat etc)
• Hur data planeras analyseras; vilka statistiska metoder för jämförelser mellan en eller
flera grupper, andra matematiska metoder, kvalitativ analys etc
-->

This is a registry-based project and all data required were collected and available prior to the project beginning. The data consists of information about trauma patients at KUH and the outcomes of morbidity & mortality conferences on selected patient cases. The next step after submitting this progress report is to begin writing the code that will import and clean data. Once this is completed, the code for data analysis will be written. The analysis will consist of using logistic regression to see if trauma severity models can predict opportunities for improvement. Initially the code will be written using scrambled data in order to minimise researcher bias. Once the code for analysis is complete, it will be run on the real-life data to obtain the results for the study. I have written a detailed time plan, including dates for deadlines, that is available for my supervisors. We will be in contact on a regular basis to discuss progress. All of my code and text will be uploaded on a regular basis, which will allow for my supervisors to see my progress.

```{r, echo=FALSE, out.width = "400px"}
knitr::include_graphics("images/timeplan.png")
```

References
==========

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->